---
title: "graphTweets - Examples"
author: "John Coene"
date: "March - 2016"
output: 
  revealjs::revealjs_presentation:
    theme: blood
    highlight: monochrome
    center: true
---

```{r, echo=FALSE, eval=TRUE, include=FALSE, message=FALSE, error=FALSE, warning=FALSE, comment=NA}
library(twitteR)
library(graphTweets)
library(igraph)
library(magrittr)
library(plotly)
```

```{r, echo=FALSE, eval=TRUE}
presi <- get(load("data/presi.RData"))
r_tw <- get(load("data/r_tw.RData"))
```

# Setup

## Install

graphTweets is available on CRAN and github.

```{r, echo=TRUE, eval=FALSE}
# from CRAN
install.packages("graphTweets")

# from github
devtools::install_github("JohnCoene/graphTweets")

# load
library(graphTweets)
```

## Functions

graphTweets only comes with *two* functions

```{r, echo=TRUE, eval=FALSE}
help(graphTweets)
```

## Data

Set yourself up with [twitteR](https://cran.r-project.org/web/packages/twitteR/index.html) or [streamR](https://cran.r-project.org/web/packages/streamR/index.html) to fetch tweets (you may of course import data in some other way).

Datasets used in these slides:

* [rstats.RData]() - `r nrow(r_tw)` tweets on #rstats.
* [presi.RData]() - `r nrow(presi)` tweets on the US Presidential 2016.

# Example 01

## search Twitter

Here I use the tiwtteR package to get 500 tweets on "#rstats". 

```{r, echo=TRUE, eval=FALSE}
library(twitteR)

#
setup_twitter_oauth(consumer_key, consumer_secret, access_token, access_secret)
r_tw <- searchTwitter("rstats", n = 500)

# to data.frame
r_tw <- twListToDF(r_tw)
```

[download]() this dataset.

## get Edges

We can now get the list of edges.

```{r, echo=TRUE, eval=TRUE}
# get edges
edges <- getEdges(r_tw, "text", "screenName")

# remove dupliactes
edges <- edges[!duplicated(edges),]
```

## igraph build

We may then use [igraph](http://igraph.org/r/) to build and plot the network. 

```{r, echo=TRUE, eval=FALSE}
library(igraph)

# build
g <- graph.data.frame(edges, directed = TRUE)

# plot
plot(g, layout = layout.fruchterman.reingold(g))
```

## igraph plot

```{r, echo=FALSE, eval=TRUE}
g <- graph.data.frame(edges, directed = TRUE)

# plot
par(bg="#262626", mar=c(1,1,1,1))
plot(g, layout = layout.fruchterman.reingold(g), vertex.color = "#DD1C1A",
     vertex.label.family = "sans",
     vertex.label.color = hsv(h = 0, s = 0, v = 0, alpha = 0.0), 
     vertex.size = log1p(degree(g)) * 2, edge.arrow.size = 0.2, 
     edge.arrow.width = 0.3, edge.width = 1,
     edge.color = hsv(h = 1, s = .59, v = .91, alpha = 0.7),
     vertex.frame.color="#262626")
```

## networkD3

You may also use the networkD3 package.

```{r, echo=TRUE, eval=FALSE}
networkD3::simpleNetwork(edges, Source = "source", Target = "target")
```

```{r, echo=FALSE, eval=TRUE}
networkD3::simpleNetwork(edges, Source = "source", Target = "target",
                         linkColour = "#262626", nodeColour = "#DD1C1A",
                         nodeClickColour = "#262626", 
                         textColour = "#580b0a") %>%
  networkD3::saveNetwork(file = 'n3.html')
```

<iframe allowtransparency="true" style="background: #262626;" src="n3.html" frameborder="0" height="410" width="100%" frameborder="0">
</iframe>

## Export

Alternatively one may use the igraph package to save the graph and import it in another software.

For instance you may want to save the graph as `graphml`.

```{r, echo=TRUE, eval=FALSE}
write.graph(g, file = "graphTweets.graphml", format = "graphml")
```

Then open it in [Gephi](https://gephi.org/).

# Example 02

## build graph

Let's use the same data set as in the previous example but this time add meta-data to the nodes.

```{r, echo=TRUE, eval=FALSE}
# graphTweets
edges <- getEdges(r_tw, source = "screenName", tweets = "text", 
                  str.length = NULL, "favoriteCount")
nodes <- getNodes(edges, source = "source", target = "target", "favoriteCount")

# replace NA with 0
nodes$retweetCount[is.na(nodes$retweetCount)] <- 0

# build graph
g <- graph.data.frame(edges, directed = TRUE, vertices = nodes)
```

```{r, echo=FALSE, eval=TRUE}
edges <- getEdges(r_tw, source = "screenName", tweets = "text", 
                  str.length = NULL, "favoriteCount")
edges$target <- as.character(edges$target)
nodes <- getNodes(edges, source = "source", target = "target", "favoriteCount")

nodes$favoriteCount[is.na(nodes$favoriteCount)] <- 0

# build graph
g <- graph.data.frame(edges, directed = TRUE, vertices = nodes)
```

## make plot

Now we can use the meta-data (`favoriteCount`) in the plot.

```{r, echo=FALSE, eval=FALSE, warning=FALSE, error=FALSE, message=FALSE}
V(g)$color <- colorRampPalette(RColorBrewer:::brewer.pal(9, "Reds"))(max(V(g)$favoriteCount))

par(bg="#262626", mar=c(1,1,1,1))
plot(g, layout = layout.fruchterman.reingold(g), vertex.color = V(g)$color,
     vertex.label.family = "sans",
     vertex.label.color = hsv(h = 0, s = 0, v = 0, alpha = 0.0), 
     vertex.size = log1p(degree(g)) * 3, edge.arrow.size = 0.2, 
     edge.arrow.width = 0.3, edge.width = 1,
     edge.color = hsv(h = 1, s = .59, v = .91, alpha = 0.7),
     vertex.frame.color="#262626")
```

## plot

```{r, echo=FALSE, eval=TRUE, warning=FALSE, error=FALSE, message=FALSE}
V(g)$color <- colorRampPalette(RColorBrewer:::brewer.pal(9, "Reds"))(max(V(g)$favoriteCount))

par(bg="#262626", mar=c(1,1,1,1))
plot(g, layout = layout.fruchterman.reingold(g), vertex.color = V(g)$color,
     vertex.label.family = "sans",
     vertex.label.color = hsv(h = 0, s = 0, v = 0, alpha = 0.0), 
     vertex.size = log1p(degree(g)) * 3, edge.arrow.size = 0.2, 
     edge.arrow.width = 0.3, edge.width = 1,
     edge.color = hsv(h = 1, s = .59, v = .91, alpha = 0.7),
     vertex.frame.color="#262626")
```

# Example 03

## Get tweets

This time we'll get tweets on the predidential race (writing this on `r Sys.Date()`)

```{r, echo=TRUE, eval=FALSE}
# search terms 2016 presidential
st <- c("hillary", "clinton", "bernie", "sanders", "donald", "trump", "jeb", 
        "bush", "ted", "cruz", "marco", "rubio", "carson")

presi <- data.frame()

for(i in 10:length(st)) {
  # tweets from the US
  tw <- searchTwitter(st[i], n = 5000, geocode = "39.554883,-99.931641,1000mi")
  
  tw <- twListToDF(tw)
  
  presi <- plyr::rbind.fill(presi, tw)
}
```

[download]() this dataset.

## get Edges

This time we pass additional variables to the `getEdges` function; geo-coordinates.

```{r, echo=TRUE, eval=TRUE}
# remove NA
presi <- presi[!is.na(presi$longitude),]

edges <- getEdges(data = presi, tweets = "text", source = "screenName", 
                  str.length = NULL, "longitude", "latitude", "retweetCount")
```

## get Nodes

Now we can add geo-coordinates as meta-data to our vertices.

```{r, echo=FALSE, eval=TRUE}
edges$source <- as.character(paste(edges$source))
edges$target <- as.character(paste(edges$target))
```

```{r, echo=TRUE, eval=TRUE}
nodes <- getNodes(edges, source = names(edges)[1], target = names(edges)[2], 
                  "longitude", "latitude", "retweetCount")
```

## build graph

Here I demonstrate with plotly as igraph cannot properly handle `nrow(nodes)` nodes.

```{r, echo=TRUE, eval=FALSE}
# join nodes and edges
edges <- dplyr::inner_join(edges[,1:2], nodes[,1:3],
                           by = c("source" = "nodes"))
edges <- dplyr::inner_join(edges, nodes[,1:3], by = c("target" = "nodes"))

names(edges)[3:ncol(edges)] <- c("lon_start", "lat_start", "lon_end",
                                 "lat_end")

# remove unknown
edges <- edges[!is.na(edges$lon_end),]

# edges frequency
edges$freq <- 1
edges <- plyr::ddply(edges, c("source", "target", "lon_start", "lat_start", 
                              "lon_end", "lat_end"),
                     plyr::summarise, freq = sum(freq))

# add id for plotly
edges$id <- 1:nrow(edges)

# plot

# map projection
geo <- list(
  scope = 'north america',
  projection = list(type = 'azimuthal equal area')
)

plot_ly(nodes, lon = longitude, lat = latitude, text = text, 
        type = 'scattergeo', locationmode = 'USA-states',
        marker = list(size = log1p(retweetCount * 3)*5, color = '#DD1C1A'),
        inherit = FALSE, hoverinfo = text) %>%
  add_trace(lon = list(lon_start, lon_end), lat = list(lat_start, lat_end),
            group = id, opacity = 1, data = edges,
            mode = 'lines', line = list(width = freq, color = '#DD1C1A'),
            type = 'scattergeo', locationmode = 'USA-states') %>%
  layout(title = 'graphTweets',
         geo = geo, showlegend = FALSE)
```

## plot

```{r, echo=FALSE, eval=TRUE}
# join nodes and edges
edges <- dplyr::inner_join(edges[,1:2], nodes[,1:3],
                           by = c("source" = "nodes"))
edges <- dplyr::inner_join(edges, nodes[,1:3], by = c("target" = "nodes"))

names(edges)[3:ncol(edges)] <- c("lon_start", "lat_start", "lon_end",
                                 "lat_end")

# remove unknown
edges <- edges[!is.na(edges$lon_end),]

# edges frequency
edges$freq <- 1
edges <- plyr::ddply(edges, c("source", "target", "lon_start", "lat_start", 
                              "lon_end", "lat_end"),
                     plyr::summarise, freq = sum(freq))

# add id
edges$id <- 1:nrow(edges)

for(i in 3:ncol(edges)){
  edges[,i] <- as.numeric(paste(edges[,i]))
}

# map projection
geo <- list(
  scope = 'north america',
  projection = list(type = 'azimuthal equal area'),
  showland = TRUE,
  landcolor = toRGB("gray75"),
  countrycolor = toRGB("#262626"),
  showocean = TRUE,
  oceancolor = "#262626",
  bgcolor = "#262626",
  caostlinecolor = "#262626",
  showsubunits = TRUE,
  subunitscolor = "#262626",
  showcountries = TRUE
)

# hover
nodes$text <- paste0(iconv(nodes$nodes, to = "UTF-8"))

# nodes retweet NA to 0
nodes$retweetCount[is.na(nodes$retweetCount)] <- 0

plot_ly(nodes, lon = longitude, lat = latitude, text = text, 
        type = 'scattergeo', locationmode = 'USA-states',
        marker = list(size = log1p(retweetCount * 3)*5, color = '#DD1C1A'),
        inherit = FALSE, hoverinfo = text) %>%
  add_trace(lon = list(lon_start, lon_end), lat = list(lat_start, lat_end),
            group = id, opacity = 1, data = edges,
            mode = 'lines', line = list(width = freq, color = '#DD1C1A'),
            type = 'scattergeo', locationmode = 'USA-states') %>%
  layout(title = '',
         geo = geo, showlegend = FALSE, paper_bgcolor = "#262626", 
         plot_bgcolor="#262626",
         width = 1000, height = 520)
```

# Outro

## Resources

If you encounter issues go to, [Github issues](https://github.com/JohnCoene/graphTweets/issues).

For the manual go on [CRAN](https://cran.r-project.org/web/packages/graphTweets/).

See my other packages and projects at [http://johncoene.github.io/projects/](http://johncoene.github.io/projects/)

# Thanks!
